```{r include = FALSE}
require(flowregime)
require(ggplot2)
require(gridExtra)
require(reshape2)
require(knitr)
opts_chunk$set(error = FALSE, warning = TRUE, message = TRUE, tidy = TRUE)
```

# Analyzing Missouri River flow data for predicting 
# Emergent Sandbar Habitat

This document outlines a time series analysis of emergent sandbar habitat 
(ESH) which attempts to identify valuable flow metrics for empirical 
predictions of ESH habitat building and loss.

```{r load-flow}
# install.packages("flowregime")
require(flowregime)
data(siouxcity)
flows = separate_flow(siouxcity, filter_parameter = 0.925)
```

```{r plot-flow}
plotdata = melt(fortify(flows), id.var = "Index", variable.name = "type", 
  value.name = "flow")
plotdata$type = factor(plotdata$type, levels = c("quickflow", "baseflow"))
ggplot(plotdata, aes(x = Index, y = flow, color = type)) + geom_line() + 
  facet_wrap(~type, ncol = 1) + xlab("") + ylab("Flow (cms)") + 
  scale_color_manual(values = c(baseflow = "blue", quickflow = "red"), 
  guide = FALSE)
```

The ESH data is contained in the geodatabase ESH_WEST.gdb; R can access file
geodatabases directly via the `rgdal` package. 

```{r load-esh}
# install.packages("rgdal")
eshw = NULL
require(rgdal)
fgdbpath = "C:/GIS workspace/EFFECTS GIS/ESH_WEST.gdb"
colnames = c("IMAGE_DATE", "TOTAL_DISCHARGE_CFS", "Class_name", "Area_ha", 
  "ESHCLASS", "a_ESH", "a_baseESH")
for(l in ogrListLayers(fgdbpath)){
   lyratt = as.data.frame(readOGR(fgdbpath, l, verbose = FALSE, 
     stringsAsFactors = FALSE))
   lyratt["LAYER_NAME"] = l
   lyratt["LAYER_DATE"] = mean(as.Date(lyratt[["IMAGE_DATE"]]))
   lyratt["LAYER_DISCHARGE"] = mean(lyratt[["TOTAL_DISCHARGE_CFS"]])
   eshw = rbind(eshw, lyratt[c(colnames, "LAYER_NAME", "LAYER_DATE", 
     "LAYER_DISCHARGE")])
}

require(dplyr)

eshw.summary = eshw %>% 
  group_by(Index = factor(LAYER_DATE), flow = round(LAYER_DISCHARGE)) %>% 
  summarize(sum_ESH = sum(a_baseESH), min_ESH = min(a_baseESH), 
    max_ESH = max(a_baseESH), mean_ESH = mean(a_baseESH), 
    sd_ESH = sd(a_baseESH))

xb = xlim(c(min(index(flows)), max(index(flows))))
flowplot = ggplot(fortify(flows), aes(x = Index)) + 
  geom_line(aes(y = baseflow), color = "blue") +
  geom_line(aes(y = baseflow + quickflow), color = "red") +
  xb + xlab("") 
  
eshplot = ggplot(eshw.summary, aes(x = as.Date(Index), y = sum_ESH, 
  size = flow)) + geom_point(color = "black", fill = "white", pch = 21) +  
  scale_size_continuous(breaks = c(10000, 20000, 30000, 40000)) +
  xb + xlab("") + ylab("Total ESH area (ha)") + theme(legend.position='bottom')
grid.arrange(flowplot, eshplot, ncol = 1) 
```
