# Verification of the IHA algorithm

This vignette verifies the `flowregime` package implementation of Indicators of
Hydrologic Alteration (IHA). The analysis of the Roanoke Rapids flow record 
described in Richter et al (1996) is repeated using both IHA v7 and the 
`IHA` function provided by `flowregime`.

```{r iha}
setwd("C:/repository")
devtools::load_all("flowregime")
data(roanokerapids)
pre.period = roanokerapids['1912-10-01/1949-09-30']
post.period = roanokerapids['1955-10-01/1991-09-30']
hpthresh = quantile(pre.period, 0.75)[[1]]
lpthresh = quantile(pre.period, 0.25)[[1]]
pre.results = IHA(pre.period, ut = hpthresh, lt = lpthresh, 
  yearstart = "10-01", yearend = "09-30", parametric = FALSE)
post.results = IHA(post.period, ut = hpthresh, lt = lpthresh, 
  yearstart = "10-01", yearend = "09-30", parametric = FALSE)
summary(pre.results)
compareIHA(pre.results, post.results)
```

The Range of Variability Approach (RVA) provided by the IHA software can also
be implemented using `flowregime`.

```{r rva}
rva.categories = build_RVA_categories(pre.results, parametric = FALSE)
rva.results = RVA(pre.results, post.results, rva.categories)

```



# References

The Nature Conservancy, 2009. Indicators of Hydrologic Alteration Version 7.1
User's Manual.

Poff, N. L. and J. V. Ward. 1989. Implications of Streamflow Variability and
Predictibility for Iotic Community Structure: a Regional Analysis of Streamflow 
Patterns. *Canadian Journal of Aquatic Sciences* 46: 1805-1818.

Richter, B. D., Baumgartner, J. V., Powell, J. and Braun, D. P. (1996), A 
Method for Assessing Hydrologic Alteration within Ecosystems. *Conservation 
Biology* 10: 1163-1174. doi: 10.1046/j.1523-1739.1996.10041163.x

Richter, B.D, J.V. Baumgartner, R. Wigington, and D.P. Braun. 1997. How much 
water does a river need? *Freshwater Biology* 37, 231-249.



no low pulses
low pulse dur
num reversals